---
title: "Fabre Museum Project: Statistical Analysis"
author: "Amélie Goubé"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: true
    toc_float: true
    theme: united
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

#Loading libraries
library(tidyverse)
library(ggpubr)
library(rstatix)
```

##1. Data Preparation 
#We import the cleaned dataset and format the grouping variable (sex).

```{r}
# Importing the cleaned file from Python
getwd()
df <- read.csv("../data/data_clean.csv", sep = ";")
# Quick chek
str(df)
```

```{r}
# Converting 'sex' into a factor to define the comparison groups
df$sex <- factor(df$sex,
                 levels = c("f", "g"),
                 labels = c("Girls", "Boys"))
```

##2. Normality Assessment
#Before choosing between a Parametric (T-test) or Non-Parametric (Wilcoxon) test, we check the distribution of each variable using the Shapiro-Wilk test.

```{r}
# List of behaviors to analyze
behavioral_vars <- c("number_of_zones_explored", "movement_autonomy", 
                     "orientation_towards_artwork", "postural_adjustment", 
                     "physical_approach", "overall_motor_engagement", 
                     "interaction_with_environment", "interaction_with_peers", 
                     "interaction_with_adults")

# Using a loop here to avoid copy-pasting the same code for every single variable
for (var in behavioral_vars) {
  cat("\n\n### Analyzing behavior:", var, "\n")
  
  
  # Checking for normality (Shapiro-Wilk test)
  # Using na.omit so the code doesn't crash if there are empty cells (NAs)
  shapiro_res <- shapiro.test(na.omit(df[[var]]))
  print(shapiro_res)
}
```

##3. Gender Comparison: Girls vs. Boys
#Since most variables are non-normal (p < 0.05) and our sample size is small (n=10),the Wilcoxon Rank Sum test is the most robust choice.

```{r}
# Creating the comparison plot and most of the data isn't normally distributed.

for (var in behavioral_vars) {
  
  # Print a header for each variable in the report
  cat("\n\n### Analyzing behavior:", var, "\n")
  
  # Normality Check (Shapiro-Wilk Test)
  # We use na.omit to handle missing data (NAs) so the test doesn't fail.
  shapiro_res <- shapiro.test(na.omit(df[[var]]))
  print(shapiro_res)
  
  # Visualizing and Comparing Groups
  # We use 'ggboxplot' from the ggpubr package.
  p <- ggboxplot(df, x = "sex", y = var,
            fill = "sex", 
            palette = "jco", 
            add = "jitter", # Jitter shows the 10 individual observation points
            title = paste("Gender effect on", var),
            subtitle = paste("Shapiro-Wilk p-value:", round(shapiro_res$p.value, 4)),
            ylab = "Observation Score", 
            xlab = "Gender (f = girls, g = boys)") +
    
    # Add the Wilcoxon test p-value directly onto the plot
    stat_compare_means(method = "wilcox.test", label.x = 1.5)
  
  print(p)
}
```

##4. Global Engagement Score
# In this section, we analyze individual behaviors and then create a composite 
# 'Global Engagement Score' to synthesize the child's overall activity.

```{r behavioral_analysis}
# Defining the variables to analyze (Quantitative & Ordinal)
variables_to_analyze <- c("number_of_zones_explored", "movement_autonomy", 
                          "orientation_towards_artwork", "postural_adjustment", 
                          "overall_motor_engagement", "interaction_with_adults")

# Running the loop for statistical analysis and visualization
# We use 'v' as a placeholder for each variable in the list
for (v in variables_to_analyze) {
  
  cat("\n\n### Statistical Analysis for:", v, "\n")
  
  # 1. Wilcoxon Rank Sum Test
  # We use rstatix:: to avoid conflicts and handle non-normal data
  test_result <- df %>% 
    rstatix::wilcox_test(as.formula(paste(v, "~ sex")))
  print(test_result)
  
  # 2. Effect Size (r)
  # Crucial for small samples (n=10) to measure the strength of the relationship
  eff_size <- df %>% 
    filter(!is.na(.data[[v]])) %>%
    rstatix::wilcox_effsize(as.formula(paste(v, "~ sex")))
  print(eff_size)
  
  # 3. Visualization with ggboxplot
  # We include jitter points to show the 10 individual observations
  p <- ggboxplot(df, x = "sex", y = v,
            fill = "sex", 
            palette = "npg", 
            add = "jitter", 
            title = paste("Comparison by Gender:", v),
            subtitle = paste("Effect Size r =", round(eff_size$effsize, 3)),
            ylab = "Observation Score", 
            xlab = "Gender Group") +
    stat_compare_means(method = "wilcox.test", label.y.npc = "top")
  
  # Display the plot in the report
  print(p)
  
} 
```

# Global Engagement Score Creation
# We merge key motor and attentional behaviors into a single indicator
# to capture the overall 'presence' of the child in the museum space.
```{r}
df$global_engagement <- rowMeans(
  df[, c("movement_autonomy", "postural_adjustment", "overall_motor_engagement")], 
  na.rm = TRUE
)

# We perform the test and SAVE it into 'global_test'
# We use rstatix:: to be safe from the error we saw earlier
global_test <- df %>% 
  rstatix::wilcox_test(global_engagement ~ sex)
print(global_test)

# Global visualization
p_global <- ggboxplot(df, x = "sex", y = "global_engagement",
            fill = "sex", palette = "jco", add = "jitter",
            title = "Global Engagement Score by Gender",
            ylab = "Mean Engagement Score", 
            xlab = "Gender") +
    stat_compare_means(method = "wilcox.test")

print(p_global)
```

```{r}

# 1. Path definition (Relative path for cross-platform compatibility)
# Defines the destination folder relative to the project root
target_path <- "results/figures/"

# Safety: Create the directory if it does not exist
if (!dir.exists(target_path)) {
  dir.create(target_path, recursive = TRUE)
}

# 2. List of variables to analyze
variables_to_analyze <- c("number_of_zones_explored", "movement_autonomy", 
                          "orientation_towards_artwork", "postural_adjustment", 
                          "overall_motor_engagement", "interaction_with_adults")

# 3. Analysis and export loop
for (v in variables_to_analyze) {
  
  # Statistical calculations (Shapiro-Wilk + Wilcoxon + Effect Size)
  shap_p <- shapiro.test(na.omit(df[[v]]))$p.value
  eff_size <- df %>% 
    filter(!is.na(.data[[v]])) %>%
    rstatix::wilcox_effsize(as.formula(paste(v, "~ sex")))
  
  # Plot creation
  p <- ggboxplot(df, x = "sex", y = v,
            fill = "sex", palette = "npg", add = "jitter",
            title = paste("Gender impact on:", v),
            subtitle = paste("Normality p:", round(shap_p, 3), " | Effect Size r:", round(eff_size$effsize, 3)),
            ylab = "Score", xlab = "Gender") +
    stat_compare_means(method = "wilcox.test", label.y.npc = "top")
  
  # Console display
  print(p)
  
  # DIRECT EXPORT to the Results folder
  # Uses the relative path defined above
  ggsave(
    filename = paste0(target_path, "boxplot_", v, ".png"),
    plot = p,
    width = 7, height = 5, dpi = 300
  )
}

# 4. Global Score and final export
df$global_engagement <- rowMeans(df[, c("movement_autonomy", "postural_adjustment", "overall_motor_engagement")], na.rm = TRUE)

p_global <- ggboxplot(df, x = "sex", y = "global_engagement",
            fill = "sex", palette = "jco", add = "jitter",
            title = "Global Engagement Score Synthesis",
            ylab = "Mean Score", xlab = "Gender") +
    stat_compare_means(method = "wilcox.test")

print(p_global)

ggsave(
  filename = paste0(target_path, "global_engagement_score.png"),
  plot = p_global,
  width = 7, height = 5, dpi = 300
)

cat("Done! All files are now in:", target_path)
```
