---
title: "Fabre Museum Project: Statistical Analysis"
author: "Amélie Goubé"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: true
    toc_float: true
    theme: united
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

#Loading libraries
library(tidyverse)
library(ggpubr)
library(rstatix)
```

# 1. Data Preparation 
We import the cleaned dataset and format the grouping variable (sex).

```{r}
# Importing the cleaned file from Python
df <- read.csv("data/data_clean.csv", sep = ";")
# Quick chek
str(df)
```

```{r}
# Converting 'sex' into a factor to define the comparison groups
df$sex <- factor(df$sex,
                 levels = c("f", "g"),
                 labels = c("Girls", "Boys"))
```

# 2. Normality Assessment
Before choosing between a Parametric (T-test) or Non-Parametric (Wilcoxon) test, we check the distribution of each variable using the Shapiro-Wilk test.

```{r}
# List of behaviors to analyze
behavioral_vars <- c("number_of_zones_explored", "movement_autonomy", 
                     "orientation_towards_artwork", "postural_adjustment", 
                     "physical_approach", "overall_motor_engagement", 
                     "interaction_with_environment", "interaction_with_peers", 
                     "interaction_with_adults")

# Using a loop here to avoid copy-pasting the same code for every single variable
for (var in behavioral_vars) {
  cat("\n\n### Analyzing behavior:", var, "\n")
  
  
  # Checking for normality (Shapiro-Wilk test)
  # Using na.omit so the code doesn't crash if there are empty cells (NAs)
  shapiro_res <- shapiro.test(na.omit(df[[var]]))
  print(shapiro_res)
}
```

# 3. Gender Comparison: Girls vs. Boys
Since most variables are non-normal (p < 0.05) and our sample size is small (n=10),the Wilcoxon Rank Sum test is the most robust choice.

```{r}
# Creating the comparison plot
for (var in behavioral_vars) {
  
  # Print a header for each variable in the report
  cat("\n\n### Analyzing behavior:", var, "\n")
  
  # Normality Check (Shapiro-Wilk Test)
  # We use na.omit to handle missing data (NAs) so the test doesn't fail.
  shapiro_res <- shapiro.test(na.omit(df[[var]]))
  print(shapiro_res)
  
  # Visualizing and Comparing Groups
p <- ggboxplot(df, x = "sex", y = var,
               fill = "sex", 
               palette = "jco", 
               add = "jitter",
               na.rm = TRUE,     
               title = paste("Gender effect on", var),
               subtitle = paste("Shapiro-Wilk p-value:", round(shapiro_res$p.value, 4)),
               ylab = "Observation Score", 
               xlab = "Gender (f = girls, g = boys)") +
    
    # Add the Wilcoxon test p-value directly onto the plot
    stat_compare_means(method = "wilcox.test", 
                       label.x = 1.5, 
                       na.rm = TRUE) 
  
  print(p)
}
```

# 4. Global Engagement Score
In this section, we analyze individual behaviors and then create a composite'Global Engagement Score' to synthesize the child's overall activity.

```{r behavioral_analysis}
# Defining the variables to analyze (Quantitative & Ordinal)
variables_to_analyze <- c("number_of_zones_explored", "movement_autonomy", 
                          "orientation_towards_artwork", "postural_adjustment", 
                          "overall_motor_engagement", "interaction_with_adults")

# Running the loop for statistical analysis and visualization
# We use 'v' as a placeholder for each variable in the list
for (v in variables_to_analyze) {
  
  cat("\n\n### Statistical Analysis for:", v, "\n")
  
  # 1. Wilcoxon Rank Sum Test
  # We use rstatix:: to avoid conflicts and handle non-normal data
  test_result <- df %>% 
    rstatix::wilcox_test(as.formula(paste(v, "~ sex")))
  print(test_result)
  
  # 2. Effect Size (r)
  # Crucial for small samples (n=10) to measure the strength of the relationship
  eff_size <- df %>% 
    filter(!is.na(.data[[v]])) %>%
    rstatix::wilcox_effsize(as.formula(paste(v, "~ sex")))
  print(eff_size)
  
  # 3. Visualization with ggboxplot
# We include jitter points to show the 10 individual observations
p <- ggboxplot(df, x = "sex", y = v,
          fill = "sex", 
          palette = "npg", 
          add = "jitter", 
          na.rm = TRUE,    
          title = paste("Comparison by Gender:", v),
          subtitle = paste("Effect Size r =", round(eff_size$effsize, 3)),
          ylab = "Observation Score", 
          xlab = "Gender Group") +
  stat_compare_means(method = "wilcox.test", 
                     label.y.npc = "top", 
                     na.rm = TRUE)   

# Display the plot in the report
print(p)
  
} 
```

# 4.2 Global Engagement Score Creation
We merge key motor and attentional behaviors into a single indicator to capture the overall 'presence' of the child in the museum space.
```{r}
df$global_engagement <- rowMeans(
  df[, c("movement_autonomy", "postural_adjustment", "overall_motor_engagement")], 
  na.rm = TRUE
)

# We perform the test and SAVE it into 'global_test'
# We use rstatix:: to be safe from the error we saw earlier
global_test <- df %>% 
  rstatix::wilcox_test(global_engagement ~ sex)
print(global_test)

# Global visualization
p_global <- ggboxplot(df, x = "sex", y = "global_engagement",
            fill = "sex", palette = "jco", add = "jitter",
            title = "Global Engagement Score by Gender",
            ylab = "Mean Engagement Score", 
            xlab = "Gender") +
    stat_compare_means(method = "wilcox.test")

print(p_global)
```

### 5. Discussion

## 5.1 Statistical Summary
The objective of this analysis was to determine if gender influences children's motor behavior and engagement at the Fabre Museum. Based on our sample of 10 observations, we can conclude the following:

* **Distribution & Test Choice:** The Shapiro-Wilk tests showed that most variables did not follow a normal distribution (p < 0.05). Consequently, we used the Wilcoxon Rank Sum test.
* **Main Results:** No statistically significant difference was found between girls and boys (p > 0.05).
* **Effect Size Insight:** We observed a moderate effect size (r ≈ 0.46) for postural adjustment, indicating a trend.

## 5.2 Scientific Interpretation
The results suggest that the museum environment acts as a "neutral" stimulus for young children. The physical interaction with the space and artworks seems to be driven more by individual curiosity than by gender-specific patterns.

## 5.3 Technical Validation
This project demonstrates a complete and reproducible data science workflow:

* **Preprocessing:** Data was cleaned and formatted using Python (Pandas).
* **Analysis:** Statistical modeling and advanced visualization were performed in R.
* **Portability:** The project uses relative paths and a standardized structure (.Rproj).

## 5.4 Limitations
The primary limitation is the sample size (N=10), which limits the power of the statistical tests. Future studies should include more subjects to confirm the trends observed.